\documentclass{article}

\usepackage[english]{babel}

% Set page size and margins
\usepackage[a4paper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{minted}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

\title{\LARGE{\textbf{Algorithm Design 21/22}}\\ \vspace{1cm} Hands On 5 - Bloom Filters}
\author{Federico Ramacciotti}
\date{}

\begin{document}
\maketitle

\section{Problem}
\begin{enumerate}
    \item Consider the Bloom filters where a single random universal hash random function $h : U \to [m]$ is employed for a set $S \subseteq U$ of keys, where $U$ is the universe of keys. 
    
    Consider its binary array $B$ of $m$ bits. Suppose that $m \geq c|S|$, for some constant $c > 1$, and that both $c$ and $|S|$ are unknown to us.
    
    Estimate the expected number of $1s$ in $B$ under a uniform choice at random of $h \in \mathcal{H}$. Is this related to $|S|$? Can we use it to estimate $|S|$?
    \item Consider $B$ and its rank function: show how to use extra $O(m)$ bits to store a space-efficient data structure that returns, for any given $i$, the following answer in constant time: 
    
    $rank(i) = \#1s \in B[1..i]$
    
    \textit{Hint}: easy to solve in extra $O(m\log m)$ bits. To get $O(m)$ bits, use prefix sums on $B$, and sample them. Use a lookup table for pieces of $B$ between any two consecutive samples.
\end{enumerate}

\section{Solution}
\subsection{Question 1}
Define the indicator variable
$$
X_i=\begin{cases}
1 & \textrm{if}\ B[i]=1\\
0 & \textrm{otherwise}
\end{cases}
$$
The probability for it to be $1$ is $Pr[X_i=1]=Pr[B[i]=1]=1-p'=1-(1-\frac{1}{m})^{|S|}$.\\
Define also the sum $Y=\sum_{i=0}^{m-1} X_i$ and find its expected value:
\begin{align*}
E[Y]&=E\left[\sum_{i=0}^{m-1} X_i\right]=\sum_{i=0}^{m-1} Pr[X_i=1]\\
&=\sum_{i=0}^{m-1} 1-\left(1-\frac{1}{m}\right)^n = m-\sum_{i=0}^{m-1}\left(1-\frac{1}{m}\right)^n\\
&=m-m\left(1-\frac{1}{m}\right)^n \cong m-me^{-\frac{n}{m}}\\
&=m\left(1-e^{-\frac{n}{m}}\right)
\end{align*}
So we get that the expected number of $1$s in $B$ is $m(1-e^{-\frac{n}{m}})$ and, as we can clearly see, it's strongly related to $|S|=n$. 
In order to estimate $|S|$, given the number of ones $n_1=E[Y]=m(1-e^{-\frac{n}{m}})$, we get $1-\frac{n_1}{m}=e^{-\frac{n}{m}}$ and, changing the sign and applying the logarithm, we obtain $n=-m\ \log(1-\frac{n_1}{m})$.

\subsection{Question 2}
\subsubsection{Baseline solution}
The baseline solution uses prefix sums in $O(m)$ time on the array $B$ to solve any query in linear time.
The space used is $O(\log m)$ bits for each integer of the array and, with an array of length $m$, the total space is $O(m\log m)$.

\subsubsection{Better solution: sampling}
In order to reduce the space used by the algorithm, we can simply sample the prefix sums array created in the baseline solution. Since the goal is to use linear space $O(m)$ with samples of size $O(\frac{\log m}{2})$, we need just $$\textrm{\# samples}*\frac{\log m}{2}=m \implies \textrm{\# samples}=\frac{2m}{\log m}$$ samples. With this solution the query time is no longer constant, because we have to find the right sample in $O(1)$ and then check the interval between two samples in $O(\frac{\log m}{2})$ time.

We can improve this solution to have a constant time per query by pre-computing a lookup table $T$ that stores the prefix sums for any possible integer of $\frac{\log m}{2}$ bits, written in binary. Thus, to get $rank(i)$, we sum the first smaller sample with the right entry in the lookup table $T$. With $m=64$ and samples of size $(\log 64)/2 = 3$, the lookup table $T$ is:
\[
    T=
    \begin{array}{ccc|ccc}
        (0 & 0 & 0)_2 & 0 & 0 & 0\\
        (0 & 0 & 1)_2 & 0 & 0 & 1\\
        (0 & 1 & 0)_2 & 0 & 1 & 1\\
        (0 & 1 & 1)_2 & 0 & 1 & 2\\
        (1 & 0 & 0)_2 & 1 & 1 & 1\\
        (1 & 0 & 1)_2 & 1 & 1 & 2\\
        (1 & 1 & 0)_2 & 1 & 2 & 2\\
        (1 & 1 & 1)_2 & 1 & 2 & 3\\
    \end{array}
\]

Since a string of $\frac{\log m}{2}$ bits can represent up to $2^{\frac{\log m}{2}}=\sqrt m$ distinct integers, the lookup table $T$ takes $\log(\frac{\log m}{2})*\sqrt m$ bits. Finally, we get that this solution takes $\sqrt{m}\frac{\log m}{2}\log(\frac{\log m}{2})=O(m)$ bits, with $O(1)$ time per query.
\end{document}